Ceph Parallel File System Evaluation Report

### 2013-08-21 cover.tex
Fix scott's name
Final (maybe) round of update for tech report version

\begin{center}
\textbf\textsc{\Large Ceph Parallel File System Evaluation Report}\\
\rule{5in}{1pt}%
\end{center}

@@

\begin{center}
Other contributors: \\
\vspace{1em}
Sarp Oral\\
Doug Fuller\\
Scotty Atchley\\
Blake Caldwell\\
James Simmons\\
Brad Settlemyer\\
Jason Hill\\
Sage Weil (Inktank Inc.)
\end{center}

>>>>

\begin{center}
\textbf{\Large Ceph Parallel File System Evaluation Report}\\
\rule{5in}{1pt}%
\end{center}

@@

\begin{center}
Other contributors: \\
\vspace{1em}
Sarp Oral\\
Doug Fuller\\
Scott Atchley\\
Blake Caldwell\\
James Simmons\\
Brad Settlemyer\\
Jason Hill\\
Sage Weil (Inktank Inc.)
\end{center}

### 2013-08-21 report.tex
Final (maybe) round of update for tech report version

National Center for Computational Sciences (NCCS), in collaboration with
Inktank Inc, has prepared a detailed performance and scalability study of
Ceph file system. Ceph is originated from Sage Weil's PhD research
at UC Santa Cruz around 2007 and it is designed to be a reliable, scalable
fault-tolerant parallel file system. Inktank is now the major developer behind
the open-source parallel file system to sheperd its development and provide
commerical support.

@@

\item Ceph has an intelligent and powerful data placement mechanism, known as
  CRUSH. The CRUSH algorithm allows a client to pre-calculate object
  placement and layout while taking into consideration failure domains and
  hierarchical storage tiers.

\item From the start, Ceph's design anticipated managing metadata and the name space
  with a cluster of metadata servers.

\item Ceph's design assumes that the system is composed of
unreliable components; fault-detection and fault-tolerance (e.g., replication) are
the norm rather than the exception. This is in line with the expectations of
Exascale computing.

@@

As part of this effort, we set up a dedicated testbed within NCCS for the Ceph
file system evaluation. The goal of our study is to investigate the feasibility
of using the Ceph for our future HPC storage deployment.
This report presents our experience, results, and observations. While evaluating
our results, please keep in mind that Ceph has not matured yet for scientific
HPC community that its code base is changing rapidly. In between releases, we
often experienced different outcomes. We will try to make clear in the report
when such changes occurred.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=4in]{figs/sfa10k}
\caption{DDN SFA10K Hardware and Host Connection Diagram}
\label{fig:ddn-sfa10k}
\end{figure}

@@

All hosts  (client and servers) were configured with Redhat 6.3 and kernel
version 3.5.1 (rhl-ceph image), Glibc 2.12 with syncfs support, locally patched.
We used the Ceph 0.55 release in the initial tests, upgraded to 0.64 and then to
0.67RC for a final round of tests.

@@

\subsection{Block I/O over Native IB}
We first established baseline performance by measuring observed block I/O
performance.
At the block-level, with each LUN configured as a RAID6 8+2 array, we had the
following results as shown in Table~\ref{tbl:block-io-baseline}.

@@

Unfortunately there was not enough time to do more detailed analysis of the network
performance. Later, we will see RADOS is performing a
maximum of 5-6 GB/s driven by four hosts.  So, it shows that we
have enough network bandwidth (i.e., IP over IB is not a bottleneck in this
case).

@@

% % Are the read and write graphs wrong or should the second sentence below be
% about write % instead of read?
The read and write performance results are summarized in
Figures~\ref{fig:xfs-read} and~\ref{fig:xfs-write}, respectively. These show
scaling behavior over the number of LUNs, which indicates that XFS can reach
the peak write performance with five SATA LUNs. Increasing number of LUNs
beyond 5, degradation happened. 
Changes to the queue depth was not observed to make any obvious difference in
the performance. For these tests, we used the \verb!xdd! benchmark with direct
I/O and a request size of 32 KB.

@@

RADOS is the Ceph object store, the foundational component for CephFS file
system. There are two types of scaling tests we are interested at the RADOS
layer:

@@

\section{Ceph File System Performance: Initial Test}

We used the synthetic IOR benchmark suite for file system level performance and
scalability test.
The particular parameter setup is show in Table \ref{tbl:ior}. Each client node
has 6 GB of physical memory, the block size is set so as to mitigate cache
effects. In addition, the test harness program issues the following commands
at the beginning of each test:

@@

\begin{itemize}
  \item The small read (4 KB transfer size) performance is almost an anomaly --
  we will investigate why it is so low compare to write performance and improved
  results in Section~\ref{sec:improve-ior}.
  \item The large read (4 MB transfer size) performance is almost half of the
  RADOS read performance.

  \item The write performance is also about half of what we can obtain from
  RADOS Bench. When number of clients reaches 8, there is a significant
  performance drop as well. 
\end{itemize}

@@

\subsection{Disable Cache Mirroring on controllers}

@@

During these tests, a trend that previously had been seen became more apparent.
During reads, there were periods of high performance followed by periods of low
performance or outright stalls that could last for up to 20 seconds at a time.
After several hours of diagnostics, Inktank observed that outstanding operations
on the clients were not being shown as outstanding on the OSDs.  This appeared
to be very similar to a problem Jim Schutt at Sandia National labs encountered
with TCP autotuning in the Linux Kernel.\footnote{\url{http://marc.info/?l=ceph-devel&m=133009796706284&w=2}}
TCP auto tuning enables TCP window scaling by default and automatically adjusts
the TCP receive window for each connection based on link conditions such as
bandwidth delay product. We have observed this will make a notable improvement
on Ceph read performance, as the results shown in
Figure~\ref{fig:rados-tcp-auto-disabled}.

@@

The initial stability issues mentioned in chapters (X,Y,Z) are fixed by
migrating from Ceph version 0.48/0.55 to 0.64, the latest stable version at the
time of writing this report.  Upgrading to the latest stable Ceph release
allowed us to run a full IOR parameter sweep for the first time since we
started evaluating the performance and scalability of the Ceph file system.
This is another sign of how much Ceph development is currently in flux.

Another fix introduced by Ceph version 0.64 was in pool creation.  The default
data pool used by previous Ceph version were set to 2x replication at the
minimum. This potentially halved the write performance. With version 0.64 we
now had the ability to set the replication level to 1, which is the preferred
value for a HPC environment like ours running on high-end and reliable storage
backend hardware (e.g. DDN SFA10K).

@@

We upgraded our test systems with kernel version 3.9 and performed a
RADOS Bench test.  The
results were extremely positive and presented in Figure~\ref{fig:rados-kernel-35vs39}.

\begin{figure}[htb]
\centering
\includegraphics[width=4in]{rados-kernel-35vs39}
\caption{RADOS bench: Kernel 3.5 vs. 3.9}
\label{fig:rados-kernel}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=4in]{ior-kernel-39}
\caption{CephFS performance with Kernel changes to 3.9, IOR with 4 MB transfer
size}
\label{fig:ior-kernel-39}
\end{figure}

@@

\item Within a single sharing mode (either shared or unique directory), stat
operations for directories and files exhibit strong linear scaling. Same strong
linear scaling is also observed for file read operations.

@@

\item Though the same saturation (or degradation) trend is not observed for file
creation operation, it is likely due to lack of workload stress on MDS.

@@

  \item Ceph is built on the assumption that the underlying hardware components
are unreliable, with little or no redundancy and failure detection capability.
This assumption is not valid for high-end HPC centers like ours. We have
disabled replication for pools, but this will take a toll somewhere in the
processing and we do not know yet if this would be significant.

>>>>

National Center for Computational Sciences (NCCS), in collaboration with
Inktank Inc, prepared this performance and scalability study of
Ceph file system. Ceph originated from Sage Weil's PhD research
at UC Santa Cruz around 2007 and it was designed to be a reliable, scalable
fault-tolerant parallel file system. Inktank is now the major developer behind
the open-source parallel file system to sheperd its development and provide
commercial support.

@@

\item Ceph has an intelligent and powerful data placement mechanism, known as
  CRUSH. The CRUSH algorithm allows a client to pre-calculate object
  placement and layout while taking into consideration of failure domains and
  hierarchical storage tiers.

\item From the start, Ceph's design anticipated managing metadata and the name space
  with a cluster of metadata servers. It utilized a dynamic subtree partitioning
  strategy to continuously adapt metadata distribution to current demands.

\item Ceph's design assumes that the system is composed of unreliable
components; fault-detection and fault-tolerance (e.g., replication) are the
norm rather than the exception. This is in line with the expectations and
future directions of Exascale computing.

@@

As part of this effort, we set up a dedicated testbed within NCCS for the Ceph
file system evaluation. The goal of our study is to investigate the
feasibility of using the Ceph for our future HPC storage deployment.  This
report presents our experience, results, and observations. While evaluating
our results, please keep in mind that Ceph is still a relatively
\textit{young} parallel file system and its code base is changing rapidly. In
between releases, we often experienced different stability and performance
outcomes.  We will try to make clear in the report when such changes occurred.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=4in]{figs/sfa10k}
\caption{DDN SFA10K hardware and host connection diagram}
\label{fig:ddn-sfa10k}
\end{figure}

@@

All hosts  (client and servers) were configured with Redhat 6.3 and kernel
version 3.5.1 initially, and later upgraded to 3.9 (rhl-ceph image), Glibc
2.12 with syncfs support, locally patched.  We used the Ceph 0.48 and 0.55
release in the initial tests, upgraded to 0.64 and then to 0.67RC for a final
round of tests.

@@

\subsection{Block I/O over Native IB}

We first established baseline performance by measuring block I/O performance.
At the block-level, with each LUN configured as a RAID6 8+2 array, we had the
following results as shown in Table~\ref{tbl:block-io-baseline}.

@@

Unfortunately there was not enough time to do more detailed analysis of the
network performance. However, as we observed later, RADOS is performing no
more than 8 GB/s driven by four server hosts. This confirms that we have
provisioned enough network bandwidth. In other words, IP over IB is not a
bottleneck in this case.

@@

% % Are the read and write graphs wrong or should the second sentence below be
% about write % instead of read?

The read and write performance results are summarized in
Figures~\ref{fig:xfs-read} and~\ref{fig:xfs-write}, respectively. These graphs
show scaling behavior over the number of LUNs, which indicates that XFS can
reach the peak write performance with just five SATA LUNs. Increasing number
of LUNs beyond 5, degradation happened.  Also, we did not observe any obvious
differences in performance when varying the queue depth parameter.  For these
tests, we used the \verb!xdd! benchmark with direct I/O and a request size of
32 KB.

@@

RADOS is a reliable distributed object store, the foundational component for
CephFS file system. There are two types of scaling tests we are interested at
the RADOS layer:

@@

\section{Ceph File System Performance: Initial Test}
\label{sec:ior-initial}
We used the synthetic IOR benchmark suite for file system level performance and
scalability test.
The particular parameter setup is show in Table \ref{tbl:ior}. Each client node
has 6 GB of physical memory, the block size is set so as to mitigate cache
effects. In addition, the test harness program issues the following commands
at the beginning of each test:

@@

  \item The small read (4 KB transfer size) performance is almost an anomaly
  -- we will investigate why it is so low compare to write performance and
  present improved results in Section~\ref{sec:improve-ior}.

  \item The large read (4 MB transfer size) performance is almost half of the
  RADOS read performance.

  \item The write performance is also about half of what we can obtain from
  RADOS Bench. When number of clients reaches 8, there is a significant
  performance drop as well.

@@

\subsection{Disable Cache Mirroring on Controllers}

@@

During these tests, a trend that previously had been seen became more apparent.
During reads, there were periods of high performance followed by periods of low
performance or outright stalls that could last for up to 20 seconds at a time.
After several hours of diagnostics, Inktank observed that outstanding operations
on the clients were not being shown as outstanding on the OSDs.  This appeared
to be very similar to a problem Jim Schutt at Sandia National labs encountered
with TCP autotuning in the Linux
kernel.\footnote{\url{http://marc.info/?l=ceph-devel&m=133009796706284&w=2}} TCP auto tuning enables TCP window scaling by default and automatically adjusts
the TCP receive window for each connection based on link conditions such as
bandwidth delay product. We have observed this will make a notable improvement
on Ceph read performance, as the results shown in
Figure~\ref{fig:rados-tcp-auto-disabled}.

@@

The initial stability issues mentioned in Section~\ref{sec:ior-initial} are
fixed by migrating from Ceph version 0.48/0.55 to 0.64, the latest stable version at the
time of writing this report.  Upgrading to the latest stable Ceph release
allowed us to run a full IOR parameter sweep for the first time since we
started evaluating the performance and scalability of the Ceph file system.
This is another sign of how much Ceph development is currently in flux.

Another fix introduced by Ceph version 0.64 was in pool creation.  The default
data pool used by previous Ceph version were set to 2x replication by mistake.
This potentially halved the write performance. With version 0.64 we explicitly
set the replication level to 1, which is the preferred value for a HPC
environment like ours running on high-end and reliable storage backend hardware
(e.g. DDN SFA10K).

@@

We upgraded our test systems with kernel version 3.9 and performed RADOS Bench
test.  The results were extremely positive and presented in
Figure~\ref{fig:rados-kernel}.

\begin{figure}[htb]
\centering
\includegraphics[width=4in]{rados-kernel-35vs39}
\caption{RADOS bench: Linux kernel version 3.5 vs. 3.9}
\label{fig:rados-kernel}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=4in]{ior-kernel-39}
\caption{CephFS performance with kernel changes to 3.9, IOR with 4 MB transfer
size}
\label{fig:ior-kernel-39}
\end{figure}

@@

\item With either shared or unique directory (\verb!-u!), stat
operations for directories and files exhibit strong linear scaling. Same strong
linear scaling is also observed for file read operations.

@@

\item Though the same saturation (or degradation) trend was not observed for
file creation operation, it is likely due to lack of workload stress on MDS.

@@

  \item Ceph is built on the assumption that the underlying hardware components
are unreliable, with little or no redundancy and failure detection capability.
This assumption is not valid for high-end HPC centers like ours. We have
disabled replication for pools, we haven't measured and quantified
processing overhead and we do not know yet if this would be significant.
